\Large Generative Dynamics of Supreme Court Citations

### 2018-08-27 Tex/paper.tex
added maggie's grant
added APSA version

\maketitle
\begin{abstract}
\noindent The significance and influence of US Supreme Court majority opinions derive in large part to opinions' roles as precedents for future opinions. A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns. We raise two limitations of existing work on Supreme Court citations. First, dyadic citations are typically aggregated to the case level before they are analyzed. Second, citations are treated as if they arise independent of one another. We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations. We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network. This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation. We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015. We find evidence for dependence processes, including reciprocity, transitivity, popularity, and activity. The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model. The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
Prepared for presentation at the ``The Enemy of My Enemy: Judicial Politics meets Network Analysis'' conference within a conference at the 2018 Annual Conference of the Southern Political Science Association. This work was supported in part by NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow.}
\end{abstract}

@@

\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
performance of the model, with some explanation of what the predictive performance means. 
I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
integrating this section into the rest of the manuscript.}

@@

The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes. The scales of covariate and dependence effects are comparable. 
%The simple linear trends represent one limitation of the current analysis. In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.

\begin{itemize}
\item adjust results/numbers in text
\end{itemize}

@@

\begin{itemize}
\item visualize Sender/receiver Issue Area Results
\item heatplot for issue area nodemix
\item visualize Sender/receiver for majority opinion writer
\item add Network effect plots for model without network statistics (add to existing plots?)
\end{itemize}

@@

\begin{itemize}
\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
\item (investigate which statistic contributes most on recall improvement)
\end{itemize}

>>>>

\maketitle
\begin{abstract}
\noindent The significance and influence of US Supreme Court majority opinions derive in large part to opinions' roles as precedents for future opinions. A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns. We raise two limitations of existing work on Supreme Court citations. First, dyadic citations are typically aggregated to the case level before they are analyzed. Second, citations are treated as if they arise independent of one another. We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations. We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network. This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation. We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015. We find evidence for dependence processes, including reciprocity, transitivity, popularity, and activity. The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model. The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. This work was supported in part by NIH, 
R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
\end{abstract}

@@

%\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
%performance of the model, with some explanation of what the predictive performance means. 
%I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
%I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
%integrating this section into the rest of the manuscript.}

@@

%The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes. The scales of covariate and dependence effects are comparable. 
%The simple linear trends represent one limitation of the current analysis. In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.

{\em {\bf Note:} The results section of this paper is incomplete and in flux. We recently discovered evidence of non-convergence in the C-TERGM results currently presented in the paper, and have changed the models and estimation settings to address this problem. Most of the substantive conclusions look like they will hold using the updated models, but we are not sure that we will be able to include the updated models in the APSA presentation.}

@@

%\begin{itemize}
%\item visualize Sender/receiver Issue Area Results
%\item heatplot for issue area nodemix
%\item visualize Sender/receiver for majority opinion writer
%\item add Network effect plots for model without network statistics (add to existing plots?)
%\end{itemize}

@@

%\begin{itemize}
%\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
%\item (investigate which statistic contributes most on recall improvement)
%\end{itemize}
